UniScatter: Unifying\\Backscatter Communications, Symbiotic Radio and Reconfigurable Intelligent Surface

### 2022-08-15 main.tex
:pencil: Update UniScatter node

\begin{document}
\title{UniScatter: Unifying\\Backscatter Communications, Symbiotic Radio and Reconfigurable Intelligent Surface}
\author{
	\IEEEauthorblockN{
		Yang~Zhao,~\IEEEmembership{Member,~IEEE,}
		and~Bruno~Clerckx,~\IEEEmembership{Fellow,~IEEE}
	}
	\thanks{
		The authors are with the Department of Electrical and Electronic Engineering, Imperial College London, London SW7 2AZ, U.K. (e-mail: \{yang.zhao18, b.clerckx\}@imperial.ac.uk).
	}
}
\maketitle

@@

	\emph{Notations:}
	Italic, bold lower-case, and bold upper-case letters denote scalars, vectors and matrices, respectively.
	$\boldsymbol{0}$ and $\boldsymbol{1}$ denote zero and one array of appropriate size, respectively.
	$\mathbb{R}_+^{x \times y}$ and $\mathbb{C}^{x \times y}$ denote the real nonnegative and complex spaces of dimension $x \times y$, respectively.
	$j$ denotes the imaginary unit.
	% $\arg(\cdot)$, $\mathrm{rank}(\cdot)$, $\mathrm{tr}(\cdot)$, $\mathrm{diag}(\cdot)$ and $\mathrm{diag}^{-1}(\cdot)$ denote the argument, rank, trace, a square matrix with input vector on main diagonal, and a vector retrieving main diagonal of input square matrix, respectively.
	$\mathrm{diag}(\cdot)$ returns a square matrix with the input vector on main diagonal.
	$(\cdot)^*$, $(\cdot)^T$, $(\cdot)^H$, $\lvert{\cdot}\rvert$, and $\lVert{\cdot}\rVert$ denote the conjugate, transpose, conjugate transpose, absolute value, and Euclidean norm operators, respectively.
	$(\cdot)^{(r)}$ and $(\cdot)^{\star}$ denote the $r$-th iterated and terminal solutions, respectively.
	The distribution of a \gls{cscg} random variable with zero mean and variance $\sigma^2$ is denoted by $\mathcal{CN}(0,\sigma^2)$, and $\sim$ means ``distributed as''.
\end{section}

\begin{section}{Scattering Principles}

@@

		\caption{Block diagram, equivalent circuit, and scatter model of a passive backscatter node. The solid and dashed vectors represent signal and energy flows. The backscatter antenna behaves as a constant power source, where the voltage $V_0$ and current $I_0$ are introduced by incident electric field $\vec{E}_{\mathrm{I}}$ and magnetic field $\vec{H}_{\mathrm{I}}$ \cite{Huang2021}.}

@@

	\begin{subsection}{Scatterer Categories}
		Scattering refers to a change of moving direction of particles after colliding with others.
		It often involves \emph{diffuse reflection} where an incident ray is redistributed at many angles.
		An electromagnetic wave is scattered when an obstruction is placed in its propagation path.
		In \gls{rf} applications such as \gls{bc}, \gls{sr} and \gls{ris}, the scatterer often includes \emph{variable-load antenna} or \emph{programmable metamaterial} with various reflection patterns \cite{Liang2022}.
		\begin{subsubsection}{Antenna-Based Scatterer}
			A typical antenna-based scatterer consists of an integrated antenna, a load-switching modulator, an energy harvester, and on-chip components (e.g., microcontroller and sensors) \cite{Dobkin2012}.
			It first receives the impinging signals, then reradiates some back to the space and dissipates the remainder.
			According to transmission line theory, the scatter-absorb tradeoff depends on the impedance matching of the load and the antenna.
			For passive scatterers without dedicated energy source, a majority of incident wave is harvested to support node operation and the reradiated part can be relatively weak \cite{Thomas2012a}.
			At the receiver, the observed signal can be further decomposed into \emph{structural mode} and \emph{antenna mode} components \cite{Hansen1989}.
			The former consistently contributes to environment multipath and is modelled within channel estimation, while the latter depends on impedance mismatch and can be used for backscatter modulation \cite{Boyer2014} and/or channel reconfiguration \cite{Liang2022}.
			For antenna-based scatterers, the reflection coefficient at load state $m$ is%
			\footnote{It corresponds to a linear scatter model where the reflection coefficient is irrelevant to incident electromagnetic field strength.}
				\Gamma_m = \frac{Z_m - Z_{\mathrm{A}}^*}{Z_m + Z_{\mathrm{A}}},
				\label{eq:reflection_pattern_antenna}
			where $Z_m$ is the load impedance at state $m$ and $Z_{\mathrm{A}}$ is the antenna input impedance.
		\end{subsubsection}
		\begin{subsubsection}{Metamaterial-Based Scatterer}
			A typical metamaterial-based scatterer consists of multiple material layers and a smart controller \cite{Wu2020}.
			The outer metamaterial layer involves numerous sub-wavelength structural units made of metallic or dielectric materials.
			The properties of unit cells can be artificially tuned for unconventional effective permittivity $\epsilon$ and permeability $\mu$, producing customizable response to the electromagnetic field \cite{Zhou2021}.
			Using epsilon-negative material (e.g., conducing rods) or mu-negative material (e.g., split-ring resonators), ideal metamaterial-based scatterers can reflect the incident waves at the boundary between free space and metamaterial without receiving them \cite{Huang2021a}.
			This mainly applies a phase shift on the reflected wave and any propagation into the metamaterial would be evanescent.
			Depending on the type of metamaterial, its equivalent impedance can be controlled by voltage (for functional materials \cite{Singh2020} and tunable diodes \cite{Costa2021}) or architecture (for resonant elements \cite{Bialkowski2008}).
			For metamaterial-based scatterers with a discrete impedance set, the reflection coefficient at metamaterial state $m$ is
				\Gamma_m = \frac{Z_m - Z_0}{Z_m + Z_0},
				\label{eq:reflection_pattern_metamaterial}
			where $Z_m$ is the equivalent impedance of metamaterial unit at state $m$ and $Z_0 = \qty{377}{\ohm}$ is the characteristic impedance of free space.
		\end{subsubsection}

@@

	\begin{subsection}{BackCom/SR: Backscatter Modulation}
		\gls{bc} and \gls{sr} commonly employ antenna-based scatterers as passive information sources.
		By \emph{frequently switching} between the available states, each backscatter node encodes self message over scattered signal and introduces backscatter uncertainty at the receiver.
		For $M$-ary \gls{qam}, reflection coefficient $\Gamma_m$ maps to the corresponding \emph{complex constellation point $c_m$} by \cite{Thomas2012a}
		\begin{equation}
			\Gamma_m = \alpha \frac{c_m}{\max_{m'} \lvert c_{m'} \rvert},
			\label{eq:backscatter_modulation}
		\end{equation}
		where $0 \le \alpha \le 1$ is the amplitude scattering ratio at the direction of interest.
	\end{subsection}

	\begin{subsection}{RIS: Channel Reconfiguration}
		\gls{ris} typically involves a planar antenna array or metasurface of numerous metamaterial units as a passive channel reconfigurator.
		By \emph{adaptively choosing} the reflection pattern, each reflecting element alters the phase of scattered signal for constructive/destructive superposition at the receiver.
		For each unit with $M$ candidate states, reflection coefficient $\Gamma_m$ maps to the corresponding \emph{phase shift $\theta_m$} by \cite{Wu2018}
		\begin{equation}
			\Gamma_m = \beta_m \exp(j \theta_m),
			\label{eq:passive_beamforming}
		\end{equation}
		where $0 \le \beta_m \le 1$ is the overall amplitude scattering ratio of state $m$.
	\end{subsection}

	% \begin{subsection}{Opportunities and Challenges}
	\begin{subsection}{Opportunities}
		We have seen conventional scattering applications as \gls{bc}, \gls{sr} and \gls{ris} using the reflection pattern \eqref{eq:reflection_pattern_antenna}, \eqref{eq:reflection_pattern_metamaterial} for either backscatter modulation \eqref{eq:backscatter_modulation} or passive beamforming \eqref{eq:passive_beamforming}.
		A natural but critical question is, for a given scatterer with pre-determined reflection pattern set, \emph{can we softly bridge backscatter modulation and passive beamforming by appropriate reflection design?}
		This question is answered in Section \ref{se:uniscatter}.
	\end{subsection}
\end{section}

\begin{section}{UniScatter}
	\begin{subsection}{UniScatter Nodes}
		% TODO compare BC, SR and IRS
		\begin{figure}[!t]
			\centering
			\subfloat[Input Distribution]{
				\resizebox{0.8\columnwidth}{!}{
					\input{assets/illustration/input_distribution.tex}
				}
			}
			\\
			\subfloat[Time Block]{
				\resizebox{0.8\columnwidth}{!}{
					\input{assets/illustration/time_block.tex}
				}
			}
			\caption{
				Input probability distribution and time block structure of \gls{sr}, \gls{ris}, and UniScatter.
				$T_s$ and $T_c$ respectively denote the primary and backscatter symbol period.
				Within channel coherence time, UniScatter nodes semi-randomly select reflection state for each backscatter block, with guidance from input probability distribution.
			}
			\label{fi:uniscatter}
		\end{figure}
		% ! Passive backscatter nodes harvest energy from and modulate information over surrounding \gls{rf} signals.
		% As shown in Fig. \subref*{fi:block_diagram}, a typical passive node consists of a scattering antenna, an energy harvester, an integrated receiver, a load-switching modulator, and on-chip components (e.g., micro-controller and sensors) \cite{Dobkin2012}.
		% Its equivalent circuit is presented in Fig. \subref*{fi:equivalent_circuit}.
		% When illuminated, the node absorbs a portion of the impinging wave for information decoding and/or energy harvesting \cite{Kim2021a}, and backscatters the remaining as \emph{structural} and \emph{antenna} components.
		% The former consistently contributes to environment multipath and can be modelled by channel estimation \cite{Boyer2014}, while the latter depends on antenna-load impedance mismatch and can be used for backscatter modulation \cite{Boyer2012} and/or channel reconfiguration \cite{Wu2021b}.
		% Fig. \subref*{fi:scatter_model} illustrates the scatter model of a node with $M$ states.

		% ? For a passive node with finite states, is it possible to unify backscatter modulation and channel reconfiguration by proper selection of reflection coefficient?
		UniScatter nodes simultaneously transmit and assist by \emph{adaptive input distribution design} based on primary and (cascaded) backscatter \gls{csi}.
		% The reflection pattern is neither fully random nor fully deterministic, but flexibly distributed to balance backscatter encoding and passive beamforming.
		% Instead of using fully random or deterministic reflection pattern over time, as shown in Fig. \ref{fi:uniscatter}, UniScatter flexibly controls input distribution of candidate reflection states to balance backscatter encoding and passive beamforming.
		% Instead of using fully random or deterministic reflection pattern over time, as shown in Fig. \ref{fi:uniscatter}, UniScatter semi-randomly selects reflection state for each backscatter block with guidance from \emph{input probability distribution $P(\Gamma_m)$.}
		% Instead of using fully random or deterministic reflection pattern over time, as shown in Fig. \ref{fi:uniscatter}, UniScatter semi-randomly selects reflection state for each backscatter block, with guidance from \emph{input probability distribution $P(\Gamma_m)$.} In other words, it flexibly controls input distribution of candidate reflection states to balance backscatter encoding and passive beamforming. \gls{sr} and \gls{ris} can be regarded as extreme cases of UniScatter, where node input distribution boils down to uniform and deterministically biased, respectively.
		Instead of using fully random or deterministic reflection pattern over time, as shown in Fig. \ref{fi:uniscatter}, UniScatter nodes semi-randomly select reflection state for each backscatter block, with \emph{guidance of input probability $P(\Gamma_m)$} for state $m$. In other words, it flexibly controls input distribution of candidate reflection states to balance backscatter encoding and passive beamforming.
		% \gls{sr} and \gls{ris} can be regarded as extreme cases of UniScatter nodes, where node input distribution boils down to uniform and deterministic, respectively.
		% flexibly controls input distribution of candidate reflection states to balance backscatter encoding and passive beamforming.
		% UniScatter semi-randomly selects reflection state for each backscatter block, with guidance from \emph{input probability distribution $P(\Gamma_m)$.}
		% % Based on \gls{csi}, UniScatter adapts the reflection state distribution to unify data transmission and channel reconfiguration.
		% \gls{sr} and \gls{ris} can be regarded as its extreme cases, where node input distribution boils down to uniform and deterministically biased, respectively.
		% Within channel coherence time, UniScatter semi-randomly selects reflection state for each backscatter block, with guidance from \emph{input probability distribution $P(\Gamma_m)$.}
		\begin{remark}
			Compared to conventional \gls{ris} literatures that optimize phase shifts under unit-module constraint, UniScatter nodes start from predefined reflection coefficients and designs their input distribution under sum-probability constraint to achieve flexible primary-backscatter tradeoff.
		\end{remark}
		% Next, we will discuss the input design in a UniScatter-enabled network.
		% consider a UniScatter-aided system to evaluate
	\end{subsection}

>>>>

\begin{document}
\title{UniScatter: Unifying\\Backscatter Communications, Symbiotic Radio and Reconfigurable Intelligent Surface}
% \title{UniScatter: Unifying Backcatter Modulation and Passive Beamforming by Adaptive Input Distribution}
\author{
	\IEEEauthorblockN{
		Yang~Zhao,~\IEEEmembership{Member,~IEEE,}
		and~Bruno~Clerckx,~\IEEEmembership{Fellow,~IEEE}
	}
	\thanks{
		The authors are with the Department of Electrical and Electronic Engineering, Imperial College London, London SW7 2AZ, U.K. (e-mail: \{yang.zhao18, b.clerckx\}@imperial.ac.uk).
	}
}
\maketitle

@@

	\emph{Notations:}
	Italic, bold lower-case, and bold upper-case letters denote scalars, vectors and matrices, respectively.
	$\boldsymbol{0}$ and $\boldsymbol{1}$ denote zero and one array of appropriate size, respectively.
	$\mathbb{R}_+^{x \times y}$ and $\mathbb{C}^{x \times y}$ denote the real nonnegative and complex spaces of dimension $x \times y$, respectively.
	$j$ denotes the imaginary unit.
	% $\arg(\cdot)$, $\mathrm{rank}(\cdot)$, $\mathrm{tr}(\cdot)$, $\mathrm{diag}(\cdot)$ and $\mathrm{diag}^{-1}(\cdot)$ denote the argument, rank, trace, a square matrix with input vector on main diagonal, and a vector retrieving main diagonal of input square matrix, respectively.
	$\mathrm{diag}(\cdot)$ returns a square matrix with the input vector on its main diagonal and zeros elsewhere.
	$(\cdot)^*$, $(\cdot)^T$, $(\cdot)^H$, $\lvert{\cdot}\rvert$, and $\lVert{\cdot}\rVert$ denote the conjugate, transpose, conjugate transpose, absolute value, and Euclidean norm operators, respectively.
	$(\cdot)^{(r)}$ and $(\cdot)^{\star}$ denote the $r$-th iterated and terminal solutions, respectively.
	The distribution of a \gls{cscg} random variable with zero mean and variance $\sigma^2$ is denoted by $\mathcal{CN}(0,\sigma^2)$, and $\sim$ means ``distributed as''.
\end{section}

\begin{section}{Scattering Principles}
	Scattering refers to a change of moving direction of particles after colliding with others.
	It often involves \emph{diffuse reflection} where an incident ray is redistributed at many angles.
	An electromagnetic wave is scattered when an obstruction is placed in its propagation path.
	In this section, we first introduce two common scatterer categories and review their operation principles, then briefly discuss how scatterers realize backscatter modulation and channel reconfiguration over reflection patterns.
	\begin{subsection}{Scatterer Categories}
		In \gls{rf} applications such as \gls{bc}, \gls{sr} and \gls{ris}, the scatterer often includes \emph{variable-load antenna} or \emph{programmable metamaterial} with various reflection patterns \cite{Liang2022}.
		\begin{subsubsection}{Antenna-Based Scatterer}
			A typical antenna-based scatterer consists of an integrated antenna, a load-switching modulator, an energy harvester, and on-chip components (e.g., microcontroller and sensors) \cite{Dobkin2012}.
			It first receives the impinging signals, then reradiates some back to the space and dissipates the remainder.
			According to transmission line theory, the scatter-absorb tradeoff depends on the impedance matching of the load and the antenna.
			For passive scatterers without dedicated energy source, a majority of incident wave is harvested to support node operation and the reradiated part can be relatively weak \cite{Thomas2012a}.
			At the receiver, the observed signal can be further decomposed into \emph{structural mode} and \emph{antenna mode} components \cite{Hansen1989}.
			The former consistently contributes to environment multipath and is modelled within channel estimation, while the latter depends on impedance mismatch and can be used for backscatter modulation \cite{Boyer2014} and/or channel reconfiguration \cite{Liang2022}.
			For antenna-based scatterers, the reflection coefficient at load state $m$ is%
			\footnote{It corresponds to a linear scatter model where the reflection coefficient is irrelevant to incident electromagnetic field strength.}
			\begin{equation}
				\Gamma_m = \frac{Z_m - Z_{\mathrm{A}}^*}{Z_m + Z_{\mathrm{A}}},
				\label{eq:reflection_pattern_antenna}
			\end{equation}
			where $Z_m$ is the load impedance at state $m$ and $Z_{\mathrm{A}}$ is the antenna input impedance.
		\end{subsubsection}
		\begin{subsubsection}{Metamaterial-Based Scatterer}
			A typical metamaterial-based scatterer consists of multiple material layers and a smart controller \cite{Wu2020}.
			The outer metamaterial layer involves numerous sub-wavelength structural units made of metallic or dielectric materials.
			The properties of unit cells can be artificially tuned for unconventional effective permittivity $\epsilon$ and permeability $\mu$, producing customizable response to the electromagnetic field \cite{Zhou2021}.
			Using epsilon-negative material (e.g., conducing rods) or mu-negative material (e.g., split-ring resonators), ideal metamaterial-based scatterers can reflect the incident waves at the boundary between free space and metamaterial without receiving them \cite{Huang2021a}.
			This mainly applies a phase shift on the reflected wave and any propagation into the metamaterial would be evanescent.
			Depending on the type of metamaterial, its equivalent impedance can be controlled by voltage (for functional materials \cite{Singh2020} and tunable diodes \cite{Costa2021}) or architecture (for resonant elements \cite{Bialkowski2008}).
			For metamaterial-based scatterers with a discrete impedance set, the reflection coefficient at metamaterial state $m$ is
			\begin{equation}
				\Gamma_m = \frac{Z_m - Z_0}{Z_m + Z_0},
				\label{eq:reflection_pattern_metamaterial}
			\end{equation}
			where $Z_m$ is the equivalent impedance of metamaterial unit at state $m$ and $Z_0 = \qty{377}{\ohm}$ is the characteristic impedance of free space.
		\end{subsubsection}
	\end{subsection}

	\begin{subsection}{BackCom/SR: Backscatter Modulation}
		\gls{bc} and \gls{sr} commonly employ antenna-based scatterers as passive information sources.
		By \emph{frequently switching} between the available states, each backscatter node encodes self message over scattered signal and introduces backscatter uncertainty at the receiver.
		For $M$-ary \gls{qam}, reflection coefficient $\Gamma_m$ maps to the corresponding \emph{complex constellation point $c_m$} by
		% \cite{Thomas2012a}
		\begin{equation}
			\Gamma_m = \alpha \frac{c_m}{\max_{m'} \lvert c_{m'} \rvert},
			\label{eq:backscatter_modulation}
		\end{equation}
		where $0 \le \alpha \le 1$ is the amplitude scattering ratio at the direction of interest.
	\end{subsection}

	\begin{subsection}{RIS: Channel Reconfiguration}
		\gls{ris} typically involves a planar antenna array or metasurface of numerous metamaterial units as a passive channel reconfigurator.
		By \emph{adaptively choosing} the reflection pattern, each reflecting element alters the phase of scattered signal for constructive/destructive superposition at the receiver.
		For each unit with $M$ candidate states, reflection coefficient $\Gamma_m$ maps to the corresponding \emph{phase shift $\theta_m$} by
		% \cite{Wu2018}
		\begin{equation}
			\Gamma_m = \beta_m \exp(j \theta_m),
			\label{eq:passive_beamforming}
		\end{equation}
		where $0 \le \beta_m \le 1$ is the overall amplitude scattering ratio of state $m$.
	\end{subsection}

	\begin{subsection}{Opportunities}
		We have seen conventional scattering applications as \gls{bc}, \gls{sr} and \gls{ris} using the reflection pattern \eqref{eq:reflection_pattern_antenna}, \eqref{eq:reflection_pattern_metamaterial} for either backscatter modulation \eqref{eq:backscatter_modulation} or passive beamforming \eqref{eq:passive_beamforming}.
		A natural but critical question is, for a given scatterer with pre-determined reflection pattern set, \emph{can we softly bridge backscatter modulation and passive beamforming by appropriate reflection design?}
		This question is answered in Section \ref{se:uniscatter}.
	\end{subsection}
\end{section}

\begin{section}{UniScatter}
	Using shared spectrum, energy and infrastructures, UniScatter is a novel passive communication/assistance protocol that generalizes \gls{bc}, \gls{sr} and \gls{ris} from the perspective of distribution.
	In particular, passive UniScatter nodes ride over an active legacy transmission in a flexible and mutualistic manner, while practical UniScatter receiver cooperatively decodes both links under the benefits of backscatter modulation and passive beamforming.
	% TODO compare BC, SR and IRS
	\begin{subsection}{UniScatter Nodes}
		Based on link priority and \gls{csi}, UniScatter nodes adapt the \emph{state input probability distribution} of antenna-based scatterers to unify and generalize backscatter modulation and passive beamforming.
		The scattering sources of \gls{bc}/\gls{sr} and reflecting elements of \gls{ris} can be viewed as its extreme cases, where the input distribution boils down to equiprobable and degenerate, respectively.
			\subfloat[Input Distribution]{
				\resizebox{0.8\columnwidth}{!}{
					\input{assets/illustration/input_distribution.tex}
				\label{fi:input_distribution}
			\\
			\subfloat[Time Block]{
				\resizebox{0.8\columnwidth}{!}{
					\input{assets/illustration/time_block.tex}
				\label{fi:time_block}
				Input probability distribution and time block structure of \gls{bc}, \gls{sr}, \gls{ris}, and UniScatter.
				$T_s$ and $T_c$ respectively denote the primary and backscatter symbol period.
				Within channel coherence time, UniScatter nodes semi-randomly select reflection state for each backscatter symbol block with the guidance of input probability distribution.
			\label{fi:uniscatter_distribution}
		As illustrated in Fig. \ref{fi:uniscatter_distribution}, instead of always using fully random or deterministic reflection pattern over time, each UniScatter node can \emph{flexibly adjust} its input distribution to balance information encoding and channel reconfiguration.
		That is, it semi-randomly select the reflection state during each backscatter symbol block with the guidance of input probability $P(\Gamma_m)$ for state $m$.
		Such an adaptive backscatter channel coding coincides with \gls{ris} when primary link is absolutely prioritized, and outperforms the uncoded equiprobable transmission of \gls{bc}/\gls{sr} when backscatter link is absolutely prioritized.
		UniScatter nodes can be implemented over readily available passive backscatter devices with energy harvester and integrated receiver \cite{Kim2021a}, while the relevant \gls{csi} can be efficiently and accurately acquired using the state-of-the-art parallel approach \cite{Jin2021a}.

@@

			\caption{Block diagram, equivalent circuit, and scatter model of a UniScatter node. The solid and dashed vectors represent signal and energy flows. The scattering antenna behaves as a constant power source, where the voltage $V_0$ and current $I_0$ are introduced by incident electric field $\vec{E}_{\mathrm{I}}$ and magnetic field $\vec{H}_{\mathrm{I}}$ \cite{Huang2021}.}
			\label{fi:uniscatter_node}

@@

		The block diagram, equivalent circuit, and scatter model of a UniScatter node are illustrated in Fig. \ref{fi:uniscatter_node}.
		A proportion of the impinging wave emitted by primary transmitter is harvested for the operation of UniScatter node, while reradiated component includes a (potentially) random phase shift due to backscatter message.
		% Next, we will discuss the input design in a UniScatter-enabled network.

@@

	\begin{subsection}{UniScatter Receiver}
		TODO
	\end{subsection}
